Regional Inequalities and economic growth; cross section and times series analysis

### 2023-11-22 MSB104_GR_1_Final_Assignment_research_article.qmd
;-)

library(tidyverse)
library(tinytex)
library(dplyr)
library(ggplot2)
library(readxl)
library(magrittr)
library(vtable)
library(lubridate)
library(dineq)
library(huxtable)
```

@@

The sructure of our study is methodically designed to facilitate this exploration. In part 1A and B we calculate GDP per capita and explore regional inequality using various descriptive statistics and visualizations such as panel data (***Assignment_1***).

In part 2A and B we use cross sectional method to examine the relationship between regional development and inequality, employing cross sectional estimation techniques for the year 2010 (***Assignment_2***).

Furthermore, exploring alternative functional forms and implementing panel estimation to analzise the relationship between regional development inequality (***Assignment_3***).

Lastly, it is worth noting potential limitations may arise due to the use of ChatGPT and ChatGPT PDF, which have been utilized as tools to gain understanding and extract knowledge from academic papers, as well as to explain econometric terms and definitions. The text generated by these tools has been cross-checked to the best of our knowledge with our textbook and Andre Seidel, lecturer and author of 'Regional Inequality, Convergence, and Its Determinants -- A View from Outer Space' (2017). Despite these precautions, there may have been instances where incorrect information was provided, and some errors could have been overlooked. The text has been meticulously rewritten to fit our style of writing, ensuring consistency and accuracy throughout our study.

@@

While correlation indicates a relationship between two variables, it doesn\'t imply causation. In econometrics, it's a common pitfall to interpret correlation as causality. Panel data helps address this by allowing researchers to control for time-invariant factors and observe changes over time, which can provide stronger evidence for causal relationships.

@@

@@

Panel data estimation is a method used in econometrics to analyse data that involves observations over multiple time periods. In the context of panel data estimation specifics of fixed-effects and random effect are used. [@lessmann2017] agrues that the fixed-effects model is a reasonable approach when the differences between countries (or regions) can be viewed as parametric shifts of the regression function. However, the random-effects model allows for time-invariant unobserved heterogeneity across regions and is more appropriate when the fixed-effects model is too restrictive [@lessmann2017].

@@

@lessmann2017 mention the use of a random-effects model to investigate the determinants of
within-country changes in inequality. The random-effects model controls for several country-level fixed factors (national income, number of regions, and area) and fixed effects for various country groups [@lessmann2017]. The advantage of the random-effects model is that the expected value of the country-specific effect is zero, which means that there is no need to apply any arbitrary data imputation procedure for the missing intercepts [@lessmann2017]. However, this approach may come at the cost of founding the predictions on a slightly biased coefficient [@lessmann2017]. The text also notes that
the major coefficient of interest is not sensitive to applying either a fixed-effects model or a random-effects model with additional country and region information [@lessmann2017].

>>>>

library(tidyverse)
library(tinytex)
library(dplyr)
library(ggplot2)
library(readxl)
library(magrittr)
library(vtable)
library(lubridate)
library(dineq)
library(huxtable)
library(plm)
```

@@

The sructure of our study is methodically designed to facilitate this exploration. In part 1A and B we calculate GDP per capita and explore regional inequality using various descriptive statistics and visualizations such as panel data (***Assignment_1***). In part 2A and B we use cross sectional method to examine the relationship between regional development and inequality, employing cross sectional estimation techniques for the year 2010 (***Assignment_2***). Furthermore, exploring alternative functional forms and implementing panel estimation to analzise the relationship between regional development inequality (***Assignment_3***).

Lastly, it is worth noting potential limitations may arise due to the use of ChatGPT and ChatGPT PDF, which have been utilized as tools to gain understanding and extract knowledge from academic papers, as well as to explain econometric terms and definitions. The text generated by these tools has been cross-checked to the best of our knowledge with our textbook and Andre Seidel, lecturer and author of 'Regional Inequality, Convergence, and Its Determinants -- A View from Outer Space' (2017). Despite these precautions, there may be instances where incorrect information was provided, and some errors could have been overlooked. The text has been meticulously rewritten to fit our style of writing, ensuring consistency and accuracy throughout our study.

@@

While correlation indicates a relationship between two variables, it doesn't imply causation. In econometrics, it's a common pitfall to interpret correlation as causality. Panel data helps address this by allowing researchers to control for time-invariant factors and observe changes over time, which can provide stronger evidence for causal relationships.

@@

Since panel data has a two-dimensional structure, with observations on multiple entities across several time periods. It enables us a more detailed and complex analyses compared to purely cross-sectional or time-series data. The data will then allow us to examine both cross-sectional (differences between subjects at a point in time) and longitudinal (changes within subjects over time) effects. For instance, this approach would be apt for studying how policy changes within a region affect its GDP and inequality levels.

@@

Panel data estimation is a method used in econometrics to analyse data that involves observations over multiple time periods. In the context of panel data estimation specifics of fixed-effects and random effect are used. [@lessmann2017] argues that the fixed-effects model is a reasonable approach when the differences between countries (or regions) can be viewed as parametric shifts of the regression function. However, the random-effects model allows for time-invariant unobserved heterogeneity across regions and is more appropriate when the fixed-effects model is too restrictive [@lessmann2017].

@@

@lessmann2017 mention the use of a random-effects model to investigate the determinants of within-country changes in inequality. The random-effects model controls for several country-level fixed factors (national income, number of regions, and area) and fixed effects for various country groups [@lessmann2017]. The advantage of the random-effects model is that the expected value of the country-specific effect is zero, which means that there is no need to apply any arbitrary data imputation procedure for the missing intercepts [@lessmann2017]. However, this approach may come at the cost of founding the predictions on a slightly biased coefficient [@lessmann2017]. @lessmann2017 also notes that the major coefficient of interest is not sensitive to applying either a fixed-effects model or a random-effects model with additional country and region information [@lessmann2017].
